---
title: "Step 2 | Sentiment_Analysis"
author: "J Argueta"
date: "4/24/2019"
output: html_document
---

Remove Objects from Environment
```{r message=FALSE}
ls()
rm(list = ls())
```

Load packages and lexicon.
```{r message=FALSE}
library(genius)#Add lyrics to a data frame
library(dplyr)
library(tidytext)
lexicon <- get_sentiments("afinn")
glimpse(lexicon)
```

### 1. Load data from step 1: top100songs.rds
```{r}
top100songs <- readRDS("top100songs.rds")
```

### 2. Data cleaning: The Genius package doesn't like special characters
```{r}
top100songsv2 <- top100songs %>% 
  mutate( Track = gsub("'", "", Track), #remove '
          Track = gsub("\\s*\\([^\\)]+\\)\\s*$", "", Track), #remove information between parenthesis
          Track = gsub("+", "", Track))
head(top100songsv2)
```

### 3. Extract lyrics for the Top 100 songs and summarise sentiment.   

This function will do the following:  
1. Pull the lyrics from the top 100 songs data frame by Artist and Track  
2. Sumarise the score of each track  

This process may take a 3 to 4 seconds per song, depending on the speed of your ISP. Also, keep in mind that not all artists may have lyrics as an output.
```{r message=FALSE, warning=FALSE}
Sentiment <- sapply(
  #X = 1:5 
  X = 1:nrow(top100songsv2)
  , FUN = function(row_num, topSongTBL){
      
      sentiment <- NA
      tryCatch({
          lyricTBL <- genius::genius_lyrics(
            artist = topSongTBL[["Artist"]][row_num]
            , song = topSongTBL[["Track"]][row_num]
          )
      
        sentiment <- lyricTBL %>%
          unnest_tokens(word, lyric) %>%
          select(word) %>%
          inner_join(lexicon) %>%
          summarise(score = sum(score))
        
        sentiment <- sentiment[[1]]
        
      }, error = function(e){
          print(paste0("Failed for song name: ", topSongTBL[["Track"]][row_num]))
      })
      
      return(sentiment)
  }
  , topSongTBL = top100songsv2
)
print(head(as.data.frame(Sentiment)))
```


### Save the final result in an RDS file top100songsSENTIMENT.rds.
```{r}
top100songsSENTIMENT <- cbind(top100songsv2, Sentiment)
saveRDS(top100songsSENTIMENT, "top100songsSENTIMENT.rds")
head(top100songsSENTIMENT)
```

