---
title: "Step 3 | Hierarchical Cluster Analysis"
author: "J Argueta"
date: "4/25/2019"
output: html_document
---

Remove Objects from Environment
```{r message=FALSE, warning=FALSE}
ls()
rm(list = ls())
```

Load packages
```{r message=FALSE, warning=FALSE}
library(tidyverse)  # data manipulation
library(stats)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```

### 1. Load Data from step 2: top100songsSENTIMENT.rds
```{r}
sentiment <- readRDS("top100songsSENTIMENT.rds")
head(sentiment)
```
### 2. Prepare data for clustering.  

Prior to clustering data, you may want to remove or estimate missing data and rescale variables for comparability.  

```{r}
sentiment1 <- sentiment %>% 
  na.omit() %>%                                   # listwise deletion of missing
  mutate(Track_Artist = paste0(Track, " by ", Artist),
         TotalStreamsScld = scale(TotalStreams),  # standardize variables
         SentimentScld = scale(Sentiment)) %>%    # standardize variables
  select(Track_Artist,TotalStreams, Sentiment, TotalStreamsScld, SentimentScld )
```

### 3. Keep only TotalStreamsScld and SentimentScld to determine number of clusters.

A plot of the within groups sum of squares by number of clusters extracted can help determine the appropriate number of clusters. The analyst looks for a bend in the plot similar to a scree test in factor analysis.  https://www.statmethods.net/about/books.html
```{r}
sentiment2 <- sentiment1 %>% 
  select(TotalStreamsScld, SentimentScld)
# Determine number of clusters
wss <- (nrow(sentiment2)-1)*sum(apply(sentiment2,2,var))
for (i in 2:15) wss[i] <- sum(kmeans(sentiment2, centers=i)$withinss)

plot(1:15, wss, type="b", xlab="Number of Clusters",ylab="Within groups sum of squares")
```


```{r}
dist_songs <- dist(sentiment2, method = "euclidean")
hc_songs <- hclust(dist_songs, method = "complete")
cluster_assignments <- cutree(hc_songs, k = 8)
cluster_assignments2 <- as.data.frame(cluster_assignments)

sentiment3 <- cbind(sentiment1, cluster_assignments2)

sentiment4 <- sentiment3 %>% 
  mutate(TotalStreamsMillions = TotalStreams/1000000)

```


### 4. Visualize the final result.

You were able to summarize 73,000 rows of data in two dimensions. PREETY COOL RIGHT!!!

Which cluster will you pick to advertise? Or which cluster will you play at the next house party? :)
```{r}
library(ggplot2)
ggplot(sentiment4, aes(x=Sentiment, y=TotalStreamsMillions, color = factor(cluster_assignments), label = Track_Artist))+
  geom_point()+
  geom_text(check_overlap = TRUE, size = 4)
```





















